---
title: "Clustering"
output:
  html_document:
#    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 120)

library(ggplot2)

library(ggfortify)

library(ranger)

library(dplyr)

library(Hmisc)

library(factoextra)

library(cluster)

library(magrittr)

library("NbClust")
```


```{r}
mydata <- read.table("~/Desktop/IMB/multivariateAnalysis/L2/Customers/Survey.csv", header=TRUE, sep=";", dec="," )
head(mydata)
```
Description:

-  ID: Customer ID.
-  x1: Customer type (1:Less than 1 year, 2:Between 1-5 years, 3:Longer than 5 years)
-  x2: Industry type (0:Restaurant, 1:Merchant)
-  x3: Firm size (0:Small firm - up to 10 employees, 1:Large firm - more than 10 employees)
-  x4: Region (0:Slo, 1:EU)
-  x5: Distribution system (0:Sold indirectly through a broker, 1:Sold directly)
-  x6: Product quality (0:Poor, 10:Excellent)
-  x7: E-commerce activities/website (0:Poor, 10:Excellent)
-  x8: Technical support (0:Poor, 10:Excellent)
-  x9: Complaint resolution (0:Poor, 10:Excellent)
- x10: Advertising (0:Poor, 10:Excellent)
- x11: Product line (0:Poor, 10:Excellent)
- x12: Sales force image (0:Poor, 10:Excellent)
- x13: Competitive pricing (0:Poor, 10:Excellent)
- x14: Warranty and claims (0:Poor, 10:Excellent)
- x15: New products (0:Poor, 10:Excellent)
- x16: Ordering and billing (0:Poor, 10:Excellent)
- x17: Price flexibility (0:Poor, 10:Excellent)
- x18: Delivery speed (0:Poor, 10:Excellent)
- x19: Customer satisfaction (0:Poor, 10:Excellent)
- x20: Likelihood of recommendation (0:Not likely, 10:Absolutely)

```{r}
mydata$CustType_F <- factor(mydata$x1, 
                            levels = c(1, 2, 3), 
                            labels = c("< 1 year", "1-5 years", "> 5 years"))

mydata$IndusType_F <- factor(mydata$x2,
                             levels = c(0, 1),
                             labels = c("Restaurant", "Merchant"))

mydata$FirmSize_F <- factor(mydata$x3, 
                            levels = c(0, 1), 
                            labels = c("Up to 10", "More than 10"))

mydata$Region_F <- factor(mydata$x4, 
                          levels = c(0, 1), 
                          labels = c("Slo", "EU"))

mydata$Distrib_F <- factor(mydata$x5, 
                           levels = c(0, 1), 
                           labels = c("Broker", "Directly"))
```

```{r}
summary(mydata[c("x6", "x8", "x12", "x15", "x17", "x18")]) #Describing clustering variables

sapply(mydata[c("x6", "x8", "x12", "x15", "x17", "x18")], FUN = var) #Calculates variance for all variables
```





```{r}
#mydata_std <- as.data.frame(scale(mydata[c("x6", "x8", "x12", "x15", "x17", "x18")])) #Saving standardized cluster variables into new data frame

sapply(mydata_std[c("x6", "x8", "x12", "x15", "x17", "x18")], FUN = var) #With standardised variance is 1
```

```{r}
mydata_std$Dissimilarity = sqrt(mydata_std$x6^2 + mydata_std$x8^2 + mydata_std$x12^2 + mydata_std$x15^2 + mydata_std$x18^2 + mydata_std$x18^2) #Finding outliers

```

```{r}
head(mydata_std[order(-mydata_std$Dissimilarity), ], 5) #Finding top 5 objects with highest value of dissimilarity
```

```{r}
print(mydata[40, ]) #Showing customer ID40
```


```{r}
mydata <- mydata[-40, ] #Removing ID40 from original data frame
mydata_std <- as.data.frame(scale(mydata[c("x6", "x8", "x12", "x15", "x17", "x18")])) 

mydata_std
```



```{r}
round(cor(mydata_std), 2)
```

```{r}
library(factoextra) 

#Finding Eudlidean distances, based on 6 Cluster variables, then saving them into object Distances

Distances <- get_dist(mydata_std, 
                      method = "euclidian")

fviz_dist(Distances) #Prikažemo matriko razdalj
```
```{r}
Distances
```



```{r}
library(factoextra) 
get_clust_tendency(mydata_std, #Hopkins statistics
                   n = nrow(mydata_std) - 1,
                   graph = FALSE) 
```

```{r}
library(dplyr)
WARD <- mydata_std %>% #Selecting variables
  get_dist(method = "euclidean") %>%  #Selecting distance
  hclust(method = "ward.D2") #Selecting algorithm         

WARD
```

```{r}
library(factoextra)
fviz_dend(WARD) #Dendrogram
```
```{r}
fviz_dend(WARD, 
          k = 3, #Number of groups
          palette = "Set1", #Colours
          rect = TRUE) #Showing groups
```


```{r}
fviz_dend(WARD, 
          k = 5, #Number of groups
          palette = "Set1", #Colours
          rect = TRUE) #Showing groups
```




```{r}
# Indices for selecting optimal number of clusters
set.seed(1)
library(NbClust)
Index <- NbClust(mydata_std,
                 distance = "euclidean",
                 method = "ward.D2", 
                 index ="all",
                 min.nc = 2, max.nc = 6) #Solution between 2 and 6 clusters

library(factoextra)
fviz_nbclust(Index,
             ggtheme = theme_linedraw())
```

```{r}
mydata$Clustering_Ward <- cutree(WARD, 
                                 k = 5) #Number of groups
head(mydata)
```


```{r}
#Showing the positions of initial leaders, used as starting point for k-means clustering

Leaders_initial <- aggregate(mydata_std, 
                             by = list(mydata$Clustering_Ward), 
                             FUN = mean)

Leaders_initial
```

```{r}
library(factoextra) 

kmeans_clu <- hkmeans(mydata_std, #Data
                      k = 5, #Number of groups
                      hc.metric = "euclidean", #Distance for hierar. clus.
                      hc.method = "ward.D2") #Alghoritm for hierar. clus.

kmeans_clu
```

```{r}
library(factoextra)
fviz_cluster(kmeans_clu, 
             palette = "Set1", 
             repel = FALSE,
             ggtheme = theme_linedraw())
```

```{r}
# Making PCA

R <- cor(mydata_std) 
round(R, 3)

library(psych)
cortest.bartlett(R, 
                 n = nrow(mydata_std))
```


```{r}
library(psych)
KMO(R)
```


```{r}
library(FactoMineR) 
components <- PCA(mydata_std, 
                  scale.unit = FALSE,
                  graph = FALSE)
```


```{r}
library(factoextra) 
get_eigenvalue(components)
```


```{r}
library(factoextra)
fviz_eig(components,
         choice = "eigenvalue",
         main = "Screeplot",
         ylab = "Eigenvalues",
         xlab = "Principal Component",
         addlabels = TRUE)
```


```{r}
library(psych)
fa.parallel(mydata_std, 
            sim = FALSE, 
            fa = "pc")
```


```{r}
library(FactoMineR)
components <- PCA(mydata_std, 
                  scale.unit = FALSE, 
                  graph = FALSE,
                  ncp = 2)
``` 


```{r}
components

print(components$var$cor)
```


```{r}
library(factoextra)
fviz_pca_var(components, 
             repel = TRUE)
```


```{r}
library(factoextra)
fviz_pca_biplot(components) 
```





```{r}
#Clustering, cont.

mydata$Clustering_kmeans <- kmeans_clu$cluster
head(mydata)
```

```{r}
#Do we have any reclassification?
table(mydata$Clustering_Ward)
table(mydata$Clustering_kmeans)
table(mydata$Clustering_Ward, mydata$Clustering_kmeans)
```

```{r}
Leaders_final <- kmeans_clu$centers
Leaders_final 
```


```{r}
#Describing clusters

Figure <- as.data.frame(Leaders_final)
Figure$ID <- 1:nrow(Figure)

library(tidyr)
Figure <- pivot_longer(Figure, cols = c("x6", "x8", "x12", "x15", "x17", "x18"))

Figure$Group <- factor(Figure$ID, 
                       levels = c(1, 2, 3, 4, 5), 
                       labels = c("1", "2", "3", "4", "5"))

Figure$NameF <- factor(Figure$name, 
                       levels = c("x6", "x8", "x12", "x15", "x17", "x18"), 
                       labels = c("Prod. value", "Tech. support", "Marketing", 
                                  "Innovativity", "Price comp.",  "Post-sale Service"))

library(ggplot2)
ggplot(Figure, aes(x = NameF, y = value)) +
  geom_hline(yintercept=0) +
  theme_linedraw() +
  geom_point(aes(shape = Group, col = Group), size=3) +
  geom_line(aes(group = ID), size=1) +
  ylab("Averages") +
  xlab("Cluster variables")+
  ylim(-2.2, 2.2)
```


```{r}
#Checking if clustering variables successfully differentiate between groups

fit <- aov(cbind(x6, x8, x12, x15, x17, x18) ~ as.factor(Clustering_kmeans), 
           data = mydata)

summary(fit)
```



```{r}
aggregate(mydata$x19, 
          by = list(mydata$Clustering_kmeans), 
          FUN = mean)
```



```{r}
aggregate(mydata$x20, 
          by = list(mydata$Clustering_kmeans), 
          FUN = mean)
```



```{r}
fit <- aov(cbind(x19, x20) ~ as.factor(Clustering_kmeans), 
           data = mydata)

summary(fit)
```

```{r}

hi_square <- chisq.test(mydata$CustType_F, as.factor(mydata$Clustering_kmeans))
hi_square

addmargins(hi_square$observed)
addmargins(round(hi_square$expected, 2)) 

round(hi_square$res, 2) 

library(DescTools)
CramerV(mydata$CustType_F, as.factor(mydata$Clustering_kmeans)) 
```


```{r}
hi_square <- chisq.test(mydata$IndusType_F, as.factor(mydata$Clustering_kmeans))
hi_square

round(hi_square$res, 2)
```

```{r}
hi_square <- chisq.test(mydata$FirmSize_F, as.factor(mydata$Clustering_kmeans))
hi_square

round(hi_square$res, 2)
```

```{r}
hi_square <- chisq.test(mydata$Region_F, as.factor(mydata$Clustering_kmeans))
hi_square

round(hi_square$res, 2)
```

```{r}
hi_square <- chisq.test(mydata$Distrib_F, as.factor(mydata$Clustering_kmeans))
hi_square

round(hi_square$res, 2)
```


```{r}
#Saving results
write.table(mydata, "/cloud/project/Lectures/L2/Customers/SurveyResults.csv", 
            row.names = FALSE, 
            sep = ";", 
            dec = ",")
```



